Reduced plasticity and relative variance in physiological rates of ectotherm populations under climate change

### 2024-06-18 ms.qmd
Few more reviewer comments dealt with

Climate change is expected to result in warmer and more variable thermal environments globally. Maintaining phenotypic variability in physiological rates and adjusting them in response to extreme temperatures (plasticity) is essential for allowing populations to adapt to climate change. Yet, incorporating both plasticity and changes in phenotypic variation when predicting the impacts of climate change on populations is often ignored. We developed effect size estimates that quantify how physiological variance changes relative to the mean as temperature increases, which we refer to as 'relative variance'. We show by meta-analysing >1900 effects from 226 species that relative variance in physiological rates decreases at higher temperatures driven by the lower proportional increase in physiological rate variance relative to the mean. Freshwater ectotherms are capable of acclimating and have the smallest reductions in relative variance. Marine organisms also showed a capacity to acclimate to higher temperatures, but capacity for plasticity traded-off with a reduction in relative variance in physiological rates at higher temperatures. Relative variance reductions were particularly pronounced for terrestrial ectotherms, and this coincided with a lack of capacity for acclimation, highlighting the vulnerability of terrestrial ectotherms to climate change. Our results show that beneficial acclimation responses in some ecosystems (marine and freshwater) may trade-off with reductions in physiological rate variance that can have important evolutionary and ecological ramifications that affect our understanding of how climate change will impact populations now and in the future.

@@

Periods of past climatic change have had disproportionate impacts on some ecosystems over others raising the question of which ecosystems will be most vulnerable to contemporary climate change. Species occupying terrestrial ecosystems are thought to be particularly vulnerable given their weak acclimation abilities and greater probability of experiencing thermal extremes that overwhelm physiological homeostasis [@Seebacher2015-wi; @Gunderson2015-hz; @hoffmann2013upper]. However, this conclusion has been questioned given that marine ectotherms have recently been shown to be closer to their upper thermal limits compared to terrestrial species [@Pinsky2019-fn]. Marine and freshwater ecosystems appear to have greater physiological acclimation capacity [e.g., @Seebacher2015-wi; @Pottier2022]. However, it is unclear if the magnitude of physiological adjustment is sufficient to compensate for potentially negative environmental effects, particularly when temperature interacts with other abiotic changes. Low oxygen availability may be a major factor influencing the vulnerability of aquatic ecosystems, whereas remaining close to thermal limits and greater water loss is expected to be a stronger constraint on physiological processes in terrestrial ectotherms [@Verberk2016]. Given the observed variation among terrestrial and aquatic ectotherms in their acclimation capacity and the different challenges associated with living in air versus water, the effect of climate warming on variation in physiological traits may differ among terrestrial and aquatic ecosystems, which may have important cascading effects on energy transfer and productivity between these ecosystems [@seebacher2023can; @barneche2021warming].

@@

A phylogeny was derived using the Open Tree of Life (OTL) with the *rotl* package in R (vers. `r utils::packageVersion("rotl")`) [@Michonneau2016-if], and plotted using *ggtree* (vers. `r utils::packageVersion("ggtree")`) [@Yu2017-wj]. We resolved all polytomies in the tree. Any missing taxa were replaced with closely related species and branch lengths were computed using Grafen's method [using power = 0.7, @Grafen1989-pr]. We used the R packages *ape* (vers. `r utils::packageVersion("ape")`) [@Paradis2019-gx] and *phytools* (vers. `r utils::packageVersion("phytools")`) [@Revell2012-vq] to prune the tree for individual analyses and calculate phylogenetic covariance (or correlation) matrices used in meta-analytic models.

@@

After quantifying levels of heterogeneity, we fit a series of multi-level meta-regression (MLMR) models to test our key questions. In all models, we included the same random effects as we used in our MLMA models. Acclimation time varied from `r range(data_wide$acclim_period, na.rm = TRUE)[1]` to `r range(data_wide$acclim_period, na.rm = TRUE)[2]` days (mean $\pm$ SD = `r mean(data_wide$acclim_period, na.rm = TRUE)` $\pm$ `r sd(data_wide$acclim_period, na.rm = TRUE)` days), and terrestrial ectotherms were acclimated for a much shorter duration (mean $\pm$ SD = `r acclim_time[3,2]` $\pm$ `r acclim_time[3,3]` days, n = `r acclim_time[3,4]`) than both freshwater (mean $\pm$ SD = `r acclim_time[1,2]` $\pm$ `r acclim_time[1,3]` days, n = `r acclim_time[1,4]`) and marine species (mean $\pm$ SD = `r acclim_time[2,2]` $\pm$ `r acclim_time[2,3]` days, n = `r acclim_time[2,4]`). Rates of acclimation have been shown to be faster for many terrestrial groups compared to aquatic organisms [e.g., amphibians and reptiles have faster rates of acclimation than fishes; See @Einum2023-lz], which would make it more likely that terrestrial ectotherms would show lower post acclimation $lnRR_{Q_{10}}$. To control for these possible differences, acclimation time was mean-centered (mean = 0) and included in all our models. As such, all estimates can be interpreted as values for an average level of acclimation time (i.e., `r mean(data_wide$acclim_period, na.rm = TRUE)` days).

We first tested the degree to which acute and acclimation $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ effects varied by habitat type (i.e., terrestrial, freshwater, and marine ecosystems). Models included an interaction between effect type (i.e., acute or acclimation) and habitat. Reduced mean $lnRR_{Q_{10_{acclimation}}}$ relative to $lnRR_{Q_{10_{acute}}}$ indicates that acclimation to thermal environments results in partial compensation of physiological rates (i.e., phenotypic plasticity), whereas no differences between $lnRR_{Q_{10_{acute}}}$ and $lnRR_{Q_{10_{acclimation}}}$ suggests organisms are not capable of physiological plasticity [@Havird2020-mj; @Seebacher2015-wi]. In contrast, a difference in $lnCVR_{Q_{10_{acclimation}}}$ relative to $lnCVR_{Q_{10_{acute}}}$ would imply that changes in between individual variation in physiological rates across 10&deg;C differ depending on whether acute or acclimation responses are measured. For both $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$, if the interaction between effect type and habitat was not supported, then we fit a model that only contained additive effects of effect type and habitat. Following on from these models, we subset each habitat type and explored how mean $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ changed across traits. Within each habitat (marine, freshwater, and terrestrial) we fit models that included an interaction between effect type (acute / acclimation) and trait category (as defined above). Variance in effects within trait categories appeared to vary depending on the trait type in question. Comparison of a model with and without heteroscedastic residual variance favored a model with heteroscedastic residual variance across trait categories ($\Delta AIC_{c}$; marine = 58, freshwater = 120, and terrestrial = 12). To ensure models converged, we did not include trait categories that had less than six effect sizes.

@@

Understanding acclimation capacity and how variation in physiological rates changes across populations and species is important for predicting the ecological and evolutionary consequences of climate change [@Chevin2017; @Chevin2010; @Bolnick2011; @bush2016incorporating; @seebacher2023can; @Sanderson2023]. Here, we show that the relative variance ($lnCVR_{Q_{10}}$) in physiological rates of ectotherms decreased across terrestrial, marine and freshwater ecosystems as temperatures increase, with the decrease being particularly pronounced in terrestrial ectotherms (~28%). These effects are expected to result in a decrease in the variability in physiological rates for marine and terrestrial ectotherms between ~4-13% under future climate change projections. Our results uncover an hitherto unrecognised dynamic where the benefits of acclimation may be counteracted by a decrease in trait variance.

@@

### ***Ecological consequences of reduced relative variation in physiological rates***
Relative variance reductions in metabolic rates and performance (e.g., sprint speed) were most strongly impacted by increased temperature. Changes in variability of traits governing energy demand could have important consequences on the flow of energy within and between populations, communities, and ecosystems [@hendry2016key; @seebacher2023can; @barneche2021warming; @Sanderson2023]. More variable populations, genetically and/or phenotypically, are predicted to be associated with broader niches, reduced intraspecific competition, increased growth rate, decreased vulnerability to environmental change and lower extinction risk [@forsman2015rethinking; @forsman2014effects; @Bolnick2011; @hendry2016key; @hart2016variation]. Maintaining intrapopulation variability in physiological rates in a warmer world may therefore be important for population resilience to climate change. For example, @kooijman1989population integrated individual variation in dynamic energy budget (DEB) models developed for *Daphnia magna*. They showed that greater between-individual variation in DEB parameters led to less extreme population fluctuations compared to scenarios with no variation [@kooijman1989population]. Our results suggest that terrestrial ectotherms will likely face greater challenges to adapting to climate change [@hoffmann2013upper; @Gunderson2015-hz; @Pottier2022], leading to greater rates of extinction and reduced productivity in terrestrial ecosystems. However, data testing the impacts of changes in variance on populations are limited to very few taxa. Future experimentation across a broader range of taxa should manipulate levels of variance to understand its ecological outcomes.

@@

Theoretical evolutionary models predict that plasticity should evolve in populations experiencing greater environmental variability (spatial or temporal), particularly when oscillations are predictable over time to make environmental cues more reliable [@Chevin2017; @Lande2009; @Chevin2010; @Nunney2016; @Murren2015; @hendry2016key; @Reed2010]. Higher spatial and temporal heterogeneity in terrestrial ecosystems [@steele2019comparison] suggests that plasticity should be more likely to evolve in terrestrial environments. However, we did not find evidence of this effect, which is consistent with other meta-analyses of heat tolerance [@barley2021limited; @Gunderson2015-hz] (which we did not consider here). Modelling correct microenvironments for over 200 species across such diverse habitats is challenging. The lack of a relationship between physiological plasticity and environmental variability and predictability could be due to the fact that ectotherms simply seek out microhabitats that make their environment quite stable [@huey2012predicting].  Therefore, it is possible that more fine-scale resolution of the habitats that organisms occupy will elucidate expected patterns. Randomly fluctuating environments have been suggested to select for reduced plasticity, particularly if plastic responses are costly and this has been demonstrated by a number of empirical studies [@Leung2020; @rescan2022experimental; @leung2023phenotypic]. For example, using seed beetles (*Callosobruchus maculatus*), @hallsson2012selection showed that experimentally evolving populations under randomly fluctuating thermal conditions did not show any increase in plasticity, but had reduced plasticity. @Leung2020 also showed reduced plasticity in morphological traits of experimentally evolving algae populations (*Dunaliella salina*) when environments were less predictable [see also @leung2023phenotypic]. While it is unclear how these findings relate directly to acclimation because they integrate multiple forms of plasticity, they do suggest that there are costs to being plastic or that the environmental signals are insufficient to trigger endocrine and epigenetic mechanisms that lead to plasticity when environments are not predictable [@Leung2020; @hendry2016key]. The fact that aquatic ectotherms did show a capacity to acclimate suggests that environmental predictability, which is expected to be higher in aquatic environments, may be more important than environmental variability in driving the evolution of plasticity.

@@

Enhanced knowledge of how variation in physiological rates vary across populations and species and the degree to which they can be adjusted in response to the environment may lead to more informed predictions about the ecological and evolutionary dynamics of natural populations [@forsman2015rethinking; @seebacher2023can; @cooke2021one; @Sanderson2023]. While we show general patterns across taxa and habitats it is important to recognise that this does not mean such patterns apply to all populations. Substantial variation in acclimation responses and changes in variance exist among populations as evidenced by wide prediction intervals. Nonetheless, changes in the relative variance in physiological rates could be better incorporated into physiological and ecological models to provide more nuanced and possibly more realistic predictions about the impacts of climate change on natural populations. While we do not yet understand the relative contribution of environmental and genetic factors to variances changes, models could better decouple how different levels of heritability with different total variance impact evolutionary and ecological predictions. Our meta-analysis now provides the opportunity to parameterise models and ensure they are better aligned with empirical findings. Nonetheless, many fascinating questions remain unanswered that will require greater focus on the consequences of changes in variance (rather than just the mean). Particularly interesting questions include: How does a reduction in physiological rate variance change energy flow across trophic levels within communities? What are the biochemical, cellular, and physiological mechanisms that underly reduced variance in physiological rates? Are reductions in variance in one trait associated with reductions in correlated traits, or do some traits increase while others decrease? How much of the reduction in variance is driven by lower levels of genetic variance? Answers to these questions will require integrative approaches that combine empirical and theoretical work across multiple levels of biological organisation but will likely provide useful advances in understanding the full consequences that climate change will have on ectotherms across all major ecosystems.

>>>>

Climate change is expected to result in warmer and more variable thermal environments globally. Maintaining phenotypic variability in physiological rates and adjusting them in response to extreme temperatures (plasticity) is essential for allowing populations to adapt to climate change. Yet, incorporating both plasticity and changes in phenotypic variation when predicting the impacts of climate change on populations is often ignored. We developed effect size estimates that quantify how physiological variance changes relative to the mean as temperature increases, which we refer to as 'relative variance'. We show by meta-analysing >1900 effects from 226 species that relative variance in physiological rates decreases at higher temperatures driven by the lower proportional increase in physiological rate variance relative to the mean. Freshwater ectotherms are capable of acclimating and have the smallest reductions in relative variance. Marine organisms also showed a capacity to acclimate to higher temperatures, but capacity for plasticity traded-off with a reduction in relative variance in physiological rates at higher temperatures. Relative variance reductions were particularly pronounced for terrestrial ectotherms, and this coincided with a lack of capacity for acclimation, highlighting the vulnerability of terrestrial ectotherms to climate change. Our results show that beneficial acclimation responses in some habitats (marine and freshwater) may trade-off with reductions in physiological rate variance that can have important evolutionary and ecological ramifications that affect our understanding of how climate change will impact populations now and in the future.

@@

Periods of past climatic change have had disproportionate impacts on some ecosystems over others raising the question of which ecosystems will be most vulnerable to contemporary climate change. Species occupying terrestrial habitats are thought to be particularly vulnerable given their weak acclimation abilities and greater probability of experiencing thermal extremes that overwhelm physiological homeostasis [@Seebacher2015-wi; @Gunderson2015-hz; @hoffmann2013upper]. However, this conclusion has been questioned given that marine ectotherms have recently been shown to be closer to their upper thermal limits compared to terrestrial species [@Pinsky2019-fn]. Marine and freshwater habitats appear to have greater physiological acclimation capacity [e.g., @Seebacher2015-wi; @Pottier2022]. However, it is unclear if the magnitude of physiological adjustment is sufficient to compensate for potentially negative environmental effects, particularly when temperature interacts with other abiotic changes. Low oxygen availability may be a major factor influencing the vulnerability of aquatic habitats, whereas remaining close to thermal limits and greater water loss is expected to be a stronger constraint on physiological processes in terrestrial ectotherms [@Verberk2016]. Given the observed variation among terrestrial and aquatic ectotherms in their acclimation capacity and the different challenges associated with living in air versus water, the effect of climate warming on variation in physiological traits may differ among terrestrial and aquatic habitats, which may have important cascading effects on energy transfer and productivity between these habitats [@seebacher2023can; @barneche2021warming].

@@

A phylogeny was derived using the Open Tree of Life (OTL) with the *rotl* package in R (vers. `r utils::packageVersion("rotl")`) [@Michonneau2016-if], and plotted using *ggtree* (vers. `r utils::packageVersion("ggtree")`) [@Yu2017-wj]. We resolved all polytomies in the tree. Any missing taxa were replaced with closely related species and branch lengths were computed using Grafen's method [using power = 0.7, @Grafen1989-pr]. Models fit using correlation matrices computed with different power (p) parameters (from 0.5 – 1.0) has nearly identical AIC. As such, we used an intermediate value of p = 0.7. We used the R packages *ape* (vers. `r utils::packageVersion("ape")`) [@Paradis2019-gx] and *phytools* (vers. `r utils::packageVersion("phytools")`) [@Revell2012-vq] to prune the tree for individual analyses and calculate phylogenetic covariance (or correlation) matrices used in meta-analytic models.

@@

After quantifying levels of heterogeneity, we fit a series of multi-level meta-regression (MLMR) models to test our key questions. In all models, we included the same random effects as we used in our MLMA models. Acclimation time varied from `r range(data_wide$acclim_period, na.rm = TRUE)[1]` to `r range(data_wide$acclim_period, na.rm = TRUE)[2]` days (mean $\pm$ SD = `r mean(data_wide$acclim_period, na.rm = TRUE)` $\pm$ `r sd(data_wide$acclim_period, na.rm = TRUE)` days), and terrestrial ectotherms were acclimated for a much shorter duration (mean $\pm$ SD = `r acclim_time[3,2]` $\pm$ `r acclim_time[3,3]` days, n = `r acclim_time[3,4]`) than both freshwater (mean $\pm$ SD = `r acclim_time[1,2]` $\pm$ `r acclim_time[1,3]` days, n = `r acclim_time[1,4]`) and marine species (mean $\pm$ SD = `r acclim_time[2,2]` $\pm$ `r acclim_time[2,3]` days, n = `r acclim_time[2,4]`). Rates of acclimation have been shown to be faster for many terrestrial groups compared to aquatic organisms [e.g., amphibians and reptiles have faster rates of acclimation than fishes; See @Einum2023-lz], which would make it more likely that terrestrial ectotherms would show lower acclimation $lnRR_{Q_{10}}$. To control for these possible differences, acclimation time was mean-centered (mean = 0) and included in all our models. As such, all estimates can be interpreted as values for an average level of acclimation time (i.e., `r mean(data_wide$acclim_period, na.rm = TRUE)` days).

We first tested the degree to which acute and acclimation $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ effects varied by habitat type (i.e., terrestrial, freshwater, and marine). Models included an interaction between effect type (i.e., acute or acclimation) and habitat. Reduced mean $lnRR_{Q_{10_{acclimation}}}$ relative to $lnRR_{Q_{10_{acute}}}$ indicates that acclimation to thermal environments results in partial compensation of physiological rates (i.e., phenotypic plasticity), whereas no differences between $lnRR_{Q_{10_{acute}}}$ and $lnRR_{Q_{10_{acclimation}}}$ suggests organisms are not capable of physiological plasticity [@Havird2020-mj; @Seebacher2015-wi]. In contrast, a difference in $lnCVR_{Q_{10_{acclimation}}}$ relative to $lnCVR_{Q_{10_{acute}}}$ would imply that changes in between individual variation in physiological rates across 10&deg;C differ depending on whether acute or acclimation responses are measured. For both $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$, if the interaction between effect type and habitat was not supported, then we fit a model that only contained additive effects of effect type and habitat. Following on from these models, we subset each habitat type and explored how mean $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ changed across traits. Within each habitat (marine, freshwater, and terrestrial) we fit models that included an interaction between effect type (acute / acclimation) and trait category (as defined above). Variance in effects within trait categories appeared to vary depending on the trait type in question. Comparison of a model with and without heteroscedastic residual variance favored a model with heteroscedastic residual variance across trait categories ($\Delta AIC_{c}$; marine = 58, freshwater = 120, and terrestrial = 12). To ensure models converged, we did not include trait categories that had less than six effect sizes.

@@

Understanding acclimation capacity and how variation in physiological rates changes across populations and species is important for predicting the ecological and evolutionary consequences of climate change [@Chevin2017; @Chevin2010; @Bolnick2011; @bush2016incorporating; @seebacher2023can; @Sanderson2023]. Here, we show that the relative variance ($lnCVR_{Q_{10}}$) in physiological rates of ectotherms decreased across terrestrial, marine and freshwater habitats as temperatures increase, with the decrease being particularly pronounced in terrestrial ectotherms (~28%). These effects are expected to result in a decrease in the variability in physiological rates for marine and terrestrial ectotherms between ~4-13% under future climate change projections. Our results uncover an hitherto unrecognised dynamic where the benefits of acclimation may be counteracted by a decrease in trait variance.

@@

### ***Ecological consequences of reduced relative variation in physiological rates***
Relative variance reductions in metabolic rates and performance (e.g., sprint speed) were most strongly impacted by increased temperature. Changes in variability of traits governing energy demand could have important consequences on the flow of energy within and between populations, communities, and ecosystems [@hendry2016key; @seebacher2023can; @barneche2021warming; @Sanderson2023]. More variable populations, genetically and/or phenotypically, are predicted to be associated with broader niches, reduced intraspecific competition, increased growth rate, decreased vulnerability to environmental change and lower extinction risk [@forsman2015rethinking; @forsman2014effects; @Bolnick2011; @hendry2016key; @hart2016variation]. Maintaining intrapopulation variability in physiological rates in a warmer world may therefore be important for population resilience to climate change. For example, @kooijman1989population integrated individual variation in dynamic energy budget (DEB) models developed for *Daphnia magna*. They showed that greater between-individual variation in DEB parameters led to less extreme population fluctuations compared to scenarios with no variation [@kooijman1989population]. Our results suggest that terrestrial ectotherms will likely face greater challenges to adapting to climate change [@hoffmann2013upper; @Gunderson2015-hz; @Pottier2022], leading to greater rates of extinction and reduced productivity in terrestrial habitats. However, data testing the impacts of changes in variance on populations are limited to very few taxa. Future experimentation across a broader range of taxa should manipulate levels of variance to understand its ecological outcomes.

@@

Theoretical evolutionary models predict that plasticity should evolve in populations experiencing greater environmental variability (spatial or temporal), particularly when oscillations are predictable over time to make environmental cues more reliable [@Chevin2017; @Lande2009; @Chevin2010; @Nunney2016; @Murren2015; @hendry2016key; @Reed2010]. Higher spatial and temporal heterogeneity in terrestrial habitats [@steele2019comparison] suggests that plasticity should be more likely to evolve in terrestrial environments. However, we did not find evidence of this effect, which is consistent with other meta-analyses of heat tolerance [@barley2021limited; @Gunderson2015-hz] (which we did not consider here). Modelling correct microenvironments for over 200 species across such diverse habitats is challenging. The lack of a relationship between physiological plasticity and environmental variability and predictability could be due to the fact that ectotherms simply seek out microhabitats that make their environment quite stable [@huey2012predicting].  Therefore, it is possible that more fine-scale resolution of the habitats that organisms occupy will elucidate expected patterns. Randomly fluctuating environments have been suggested to select for reduced plasticity, particularly if plastic responses are costly and this has been demonstrated by a number of empirical studies [@Leung2020; @rescan2022experimental; @leung2023phenotypic]. For example, using seed beetles (*Callosobruchus maculatus*), @hallsson2012selection showed that experimentally evolving populations under randomly fluctuating thermal conditions did not show any increase in plasticity, but had reduced plasticity. @Leung2020 also showed reduced plasticity in morphological traits of experimentally evolving algae populations (*Dunaliella salina*) when environments were less predictable [see also @leung2023phenotypic]. While it is unclear how these findings relate directly to acclimation because they integrate multiple forms of plasticity, they do suggest that there are costs to being plastic or that the environmental signals are insufficient to trigger endocrine and epigenetic mechanisms that lead to plasticity when environments are not predictable [@Leung2020; @hendry2016key]. The fact that aquatic ectotherms did show a capacity to acclimate suggests that environmental predictability, which is expected to be higher in aquatic environments, may be more important than environmental variability in driving the evolution of plasticity.

@@

Enhanced knowledge of how variation in physiological rates vary across populations and species and the degree to which they can be adjusted in response to the environment may lead to more informed predictions about the ecological and evolutionary dynamics of natural populations [@forsman2015rethinking; @seebacher2023can; @cooke2021one; @Sanderson2023]. While we show general patterns across taxa and habitats it is important to recognise that this does not mean such patterns apply to all populations. Substantial variation in acclimation responses and changes in variance exist among populations as evidenced by wide prediction intervals. Nonetheless, changes in the relative variance in physiological rates could be better incorporated into physiological and ecological models to provide more nuanced and possibly more realistic predictions about the impacts of climate change on natural populations. While we do not yet understand the relative contribution of environmental and genetic factors to variances changes, models could better decouple how different levels of heritability with different total variance impact evolutionary and ecological predictions. Our meta-analysis now provides the opportunity to parameterise models and ensure they are better aligned with empirical findings. Nonetheless, many fascinating questions remain unanswered that will require greater focus on the consequences of changes in variance (rather than just the mean). Particularly interesting questions include: How does a reduction in physiological rate variance change energy flow across trophic levels within communities? What are the biochemical, cellular, and physiological mechanisms that underly reduced variance in physiological rates? Are reductions in variance in one trait associated with reductions in correlated traits, or do some traits increase while others decrease? How much of the reduction in variance is driven by lower levels of genetic variance? Answers to these questions will require integrative approaches that combine empirical and theoretical work across multiple levels of biological organisation but will likely provide useful advances in understanding the full consequences that climate change will have on ectotherms across all major habitats.
